<<<<<<< Updated upstream
---
title: "NYPD Motor Vehicle Collisions Analysis"
output:
  html_document:
    toc: true
    theme: united
---

# Data summary

```{r}
#Imports
library(ggmap)
library(cowplot)
library(gganimate)
library(animation)
library(ggplot2)
library(zoo) 
library(dplyr)
library(plyr)
library(stringr)

# Google Maps Style
actual_style <- "https://maps.googleapis.com/maps/api/staticmap?key=AIzaSyBmeGRbELjcBo0FUmB5LTK0ZUcBNKrcdwQ&center=40.710214438068604,-73.9838008730469&zoom=12&format=png&maptype=roadmap&style=saturation:-100%7Cgamma:0.5&style=feature:administrative%7Celement:geometry%7Cvisibility:off&style=feature:administrative.land_parcel%7Cvisibility:off&style=feature:administrative.neighborhood%7Cvisibility:off&style=feature:poi%7Cvisibility:off&style=feature:poi%7Celement:labels.text%7Cvisibility:off&style=feature:road%7Celement:labels%7Cvisibility:off&style=feature:road%7Celement:labels.icon%7Cvisibility:off&style=feature:road.arterial%7Celement:labels%7Cvisibility:off&style=feature:road.highway%7Celement:labels%7Cvisibility:off&style=feature:road.local%7Cvisibility:off&style=feature:transit%7Cvisibility:off&style=feature:water%7Celement:labels.text%7Cvisibility:off&size=480x360"

# Specify columnn type to soeed up loading
columnTypes = c("DATE"="character", "TIME"="factor", "ZIP.CODE"="integer", "LATITUDE"="double", "LONGITUDE"="double",
  "LOCATION"="factor", "ON.STREET.NAME"="factor", "CROSS.STREET.NAME"="factor", "OFF.STREET.NAME"="factor",
  "NUMBER.OF.PERSONS.INJURED"="integer", "NUMBER.OF.PERSONS.KILLED"="integer",
  "NUMBER.OF.PEDESTRIANS.INJURED"="integer","NUMBER.OF.PEDESTRIANS.KILLED"="integer", 
  "NUMBER.OF.CYCLIST.INJURED"="integer", "NUMBER.OF.CYCLIST.KILLED"="integer", "NUMBER.OF.MOTORIST.INJURED"="integer", 
  "NUMBER.OF.MOTORIST.KILLED"="integer", "CONTRIBUTING.FACTOR.VEHICLE.1"="factor", 
  "CONTRIBUTING.FACTOR.VEHICLE.2"="factor", "CONTRIBUTING.FACTOR.VEHICLE.3"="factor", 
  "CONTRIBUTING.FACTOR.VEHICLE.4"="factor", "CONTRIBUTING.FACTOR.VEHICLE.5"="factor", "UNIQUE.KEY"="factor", 
  "VEHICLE.TYPE.CODE.1"="integer", "VEHICLE.TYPE.CODE.1"="factor", "VEHICLE.TYPE.CODE.2"="factor", 
  "VEHICLE.TYPE.CODE.3"="factor", "VEHICLE.TYPE.CODE.4"="factor", "VEHICLE.TYPE.CODE.5"="factor")

#Load data set
collisions <- read.csv(file = "NYPD_Motor_Vehicle_Collisions.csv", header =TRUE, 
                      colClasses = columnTypes)

collisions$DATE = as.Date(collisions$DATE, format="%m/%d/%Y")
```


# Amount of fatal collisions
```{r}
nrow(collisions[collisions$NUMBER.OF.PERSONS.KILLED>0,])
```



# Plot the known accident locations

```{r}
# only retain the data point having lat/long info
collisionsComplete = collisions[complete.cases(collisions$LONGITUDE) & complete.cases(collisions$LATITUDE),]

# Get a map
#map = get_map(location = "New York city", source="osm", maptype="roadmap", zoom = "auto", color = "color")
map = get_googlemap(center="New York City", maptype="terrain", zoom=10, color="color", size = c(640, 420), style = actual_style)

#ggmap(map, extent = "device") +
 #   geom_point(data=collisionsComplete, aes(x = LONGITUDE, y = LATITUDE),
  #             color = "#c43400", size = 1, alpha=0.5)
```


# Collisions heatmap

```{r}
# Draw heat map
ggmap(map, extent = "device") +
  geom_density2d(data = collisionsComplete, aes(x = LONGITUDE, y = LATITUDE), size = 0.3) +stat_density2d(data = collisionsComplete, aes(x = LONGITUDE, y = LATITUDE, fill = ..level.., alpha = ..level..), size = 0.01, bins = 16, geom = "polygon") +scale_fill_gradient(low = "#ff6c6c", high = "#930000") + scale_alpha(range = c(0.2, 0.5), guide = FALSE)
```


# Mortal collisions heatmap

```{r}
# Draw the heat map
mortalCollisions = collisions[collisions$NUMBER.OF.PERSONS.KILLED > 0 & complete.cases(collisions$LONGITUDE) & complete.cases(collisions$LATITUDE), ]

ggmap(map, extent = "device") +
  geom_density2d(data = mortalCollisions, aes(x = LONGITUDE, y = LATITUDE), size = 0.3) +
  stat_density2d(data = mortalCollisions, aes(x = LONGITUDE, y = LATITUDE, fill = ..level.., alpha = ..level..), size = 0.01, bins = 16, geom = "polygon") +
  scale_fill_gradient(low = "green", high = "red") +
  scale_alpha(range = c(0, 0.3), guide = FALSE)
```


# Animated map of every collision over time

```{r}

# Retain mortal collisions in Manahattan only


animatedMap <- ggmap(map, extent = "device") +
  geom_point(data = collisions[as.Date("2012-06-01") <collisions$DATE & 
                                       collisions$DATE<as.Date("2012-08-01"), ],
             aes(x = LONGITUDE, y =LATITUDE, frame=DATE, cumulative=TRUE),
             color = "#c43400", size = 1, alpha=0.3) +
    theme(plot.title = element_text(hjust = 0.5))

animation::ani.options(interval=0.5)
gganimate(animatedMap,ani.width = 750, ani.height = 750)
```

```{r}
gganimate(animatedMap, filename = "accidents.gif", ani.width = 640, ani.height = 420)
```


# Animated map of deadly collisions over time

```{r}

# Retain mortal collisions in Manahattan only

mortalCollisions <- transform(mortalCollisions, YEARMONTH = as.yearmon(DATE)) 

animatedMap <- ggmap(map) +
    geom_point(data = mortalCollisions,
             aes(x = LONGITUDE, y =LATITUDE, frame=YEARMONTH, cumulative=TRUE),
             color = "#c43400", size = 2, alpha=0.3) +
    theme(plot.title = element_text(hjust = 0.5))

animation::ani.options(interval=0.5)
gganimate(animatedMap,ani.width = 750, ani.height = 750)
```


# Animation of deadly collisions in Manhattan over time

Let's look a bit closer at the deadly collisions in Manhattan.

```{r}
manhattan = get_googlemap(center="Manhattan", maptype="terrain", zoom=11, color="color", size = c(640, 420))

# Retain mortal collisions in Manahattan only
mortalCollisionsManhattan = collisionsComplete[collisionsComplete$BOROUGH =="MANHATTAN" & collisionsComplete$NUMBER.OF.PERSONS.KILLED>0,]


mortalCollisionsManhattan <- transform(mortalCollisionsManhattan, YEARMONTH = as.yearmon(DATE)) 

animatedMap <- ggmap(manhattan) +
    geom_point(data = mortalCollisionsManhattan,
             aes(x = LONGITUDE, y =LATITUDE, frame=YEARMONTH, cumulative=TRUE),
             color = "#c43400", size = 2, alpha=0.3) +
    theme(plot.title = element_text(hjust = 0.5))

animation::ani.options(interval=1)
gganimate(animatedMap,ani.width = 750, ani.height = 750)
```


# Collisions per borough

```{r}
boroughs = levels(collisions$BOROUGH) # Get boroughs' name

frequency = c(0,0,0,0,0,0) # Initialize frequency counts
i=1
for (borough in boroughs){ # Get accident frequency per borough
  frequency[i] = nrow(collisions[collisions$BOROUGH==borough,])
  i =i+1
}

newOrdering = sort(frequency, decreasing=TRUE, index.return=TRUE)$ix # index of boroughs in decreasing order

boroughs = factor(collisions$BOROUGH,boroughs[newOrdering]) # Order borough in decreasing order
levels(boroughs)[1] = "Unknown" # Replace "" by Unknown

# Plot bar plot
boroughsDataframe <- data.frame(boroughs)
ggplot(data=boroughsDataframe, aes(boroughs, fill=boroughs)) +
  geom_bar()+ggtitle("Count of collisions per borough")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Mortal collisions per borough

```{r}
mortalAccidents = collisions[collisions$NUMBER.OF.PERSONS.KILLED > 0, ]

boroughs = levels(mortalAccidents$BOROUGH) # Get boroughs' name

frequency = c(0,0,0,0,0,0) # Initialize frequency counts
i=1
for (borough in boroughs){ # Get accident frequency per borough
  frequency[i] = nrow(mortalAccidents[mortalAccidents$BOROUGH==borough,])
  i =i+1
}

newOrdering = sort(frequency, decreasing=TRUE, index.return=TRUE)$ix # index of boroughs in decreasing order

boroughs = factor(mortalAccidents$BOROUGH,boroughs[newOrdering]) # Order borough in decreasing order
levels(boroughs)[1] = "Unknown" # Replace "" by Unknown

# Plot bar plot
boroughsDataframe <- data.frame(boroughs)
ggplot(data=boroughsDataframe, aes(boroughs, fill=boroughs)) +
  geom_bar()+ggtitle("Count of mortal collisions per borough") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Most common causes of collisions

```{r}

# Concatenate every causes together
contributingFactors = unlist(list(collisions$CONTRIBUTING.FACTOR.VEHICLE.1,
                                  collisions$CONTRIBUTING.FACTOR.VEHICLE.2, 
                                  collisions$CONTRIBUTING.FACTOR.VEHICLE.3, 
                                  collisions$CONTRIBUTING.FACTOR.VEHICLE.4, 
                                  collisions$CONTRIBUTING.FACTOR.VEHICLE.5))

# Remove unknown causes
contributingFactors = as.character(contributingFactors)
contributingFactors = contributingFactors[contributingFactors != "Unspecified" & contributingFactors != ""]
contributingFactors = as.factor(contributingFactors)


################
  #Create a custom color scale for the contributing factors only
  library(RColorBrewer)
  colourCount <- length(levels(contributingFactors))
  getPalette = colorRampPalette(brewer.pal(9, "Blues")) #choose other set to get another color scheme
  #myColors = getPalette(colourCount)
  myColors = getPalette(30)
  myColors = unlist(list(rep_len("#000000", (colourCount - 30)), myColors))
 
  #names(myColors) <- levels(contributingFactors)
  colScale <-  scale_fill_manual(name = "grp", values = myColors)
################
  
mostCommonCauses = plyr::count(contributingFactors) # Get frequency counts
names(myColors) <- mostCommonCauses$x[order(mostCommonCauses$freq)] # Assign the red color in descending color
mostCommonCauses=mostCommonCauses[order(-mostCommonCauses$freq),]


# Create dataframe with top ten causes and frequencies
contributingFactorsDataframe <- data.frame("causes"=as.character(mostCommonCauses$x[1:10]), "frequency"=
                                             mostCommonCauses$freq[1:10]) #keep top 10 only

contributingFactorsDataframe$causes = factor(contributingFactorsDataframe$causes, 
                                             levels = contributingFactorsDataframe$causes)

p <- ggplot(data=contributingFactorsDataframe, aes(x=causes, y=frequency, fill=causes)) +
  geom_bar(stat = "identity")+ggtitle("Most common known causes of collisions in NYC")+
  theme(axis.text.x = element_text(angle=45, hjust = 1, size=12), legend.position="None")+
  scale_fill_manual(values = myColors)+
  ylab("Number of collisions")+xlab(NULL)
p
```


# Most common causes of collisions per borough

```{r}
#par(mfrow=c(3,3))
for(borough in levels(collisions$BOROUGH)){
  collisionsInBorough = collisions[collisions$BOROUGH==borough, ] # Collision in current borough
  # Concatenate every causes together
  contributingFactors = unlist(list(collisionsInBorough$CONTRIBUTING.FACTOR.VEHICLE.1, 
                                    collisionsInBorough$CONTRIBUTING.FACTOR.VEHICLE.2, 
                                    collisionsInBorough$CONTRIBUTING.FACTOR.VEHICLE.3,
                                    collisionsInBorough$CONTRIBUTING.FACTOR.VEHICLE.4, 
                                    collisionsInBorough$CONTRIBUTING.FACTOR.VEHICLE.5))
  
  # Remove unknown causes
  contributingFactors = as.character(contributingFactors)
  contributingFactors = contributingFactors[contributingFactors != "Unspecified" & contributingFactors != ""]
  contributingFactors = as.factor(contributingFactors)
  contributingFactors
  
  mostCommonCauses = plyr::count(contributingFactors) # Get frequency counts
  mostCommonCauses=mostCommonCauses[order(-mostCommonCauses$freq),]
  
  # Create dataframe with top ten causes and frequencies
  contributingFactorsDataframe <- data.frame("causes"=as.character(mostCommonCauses$x[1:10]), "frequency"=
                                               mostCommonCauses$freq[1:10]) #keep top 10 only
  
  contributingFactorsDataframe$causes = factor(contributingFactorsDataframe$causes, levels =
                                                 contributingFactorsDataframe$causes)
  plotTitle = paste("Most common known causes of collisions in", borough)
  print(ggplot(data=contributingFactorsDataframe, aes(x=causes, y=frequency, fill=causes)) +
          geom_bar(stat = "identity")+
          ggtitle(plotTitle)+
          theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1)))
}
```


# Most common causes of mortal collisions

Note: this does not contain the accidents whose cause remained unknown. Look how Traffic control disregard jumped to the second place now while it was the 10th cause overall. Alcohol and unsafe speed are now 5th and 6th while they were not in the top 10 before. Passenger distraction also appears now.
```{r}
mortalCollisions = collisions[collisions$NUMBER.OF.PERSONS.KILLED > 0, ]

# Concatenate every causes together
contributingFactors = unlist(list(mortalCollisions$CONTRIBUTING.FACTOR.VEHICLE.1,
                                  mortalCollisions$CONTRIBUTING.FACTOR.VEHICLE.2, 
                                  mortalCollisions$CONTRIBUTING.FACTOR.VEHICLE.3, 
                                  mortalCollisions$CONTRIBUTING.FACTOR.VEHICLE.4, 
                                  mortalCollisions$CONTRIBUTING.FACTOR.VEHICLE.5))

# Remove unknown causes
contributingFactors = as.character(contributingFactors)
contributingFactors = contributingFactors[contributingFactors != "Unspecified" & contributingFactors != ""]
contributingFactors = as.factor(contributingFactors)

mostCommonCauses = plyr::count(contributingFactors) # Get frequency counts
mostCommonCauses=mostCommonCauses[order(-mostCommonCauses$freq),]

# Create dataframe with top ten causes and frequencies
contributingFactorsDataframe <- data.frame("causes"=as.character(mostCommonCauses$x[1:10]), "frequency"=
                                             mostCommonCauses$freq[1:10]) #keep top 10 only

contributingFactorsDataframe$causes = factor(contributingFactorsDataframe$causes, levels = contributingFactorsDataframe$causes)
p <- ggplot(data=contributingFactorsDataframe, aes(x=causes, y=frequency, fill=causes))+
  geom_bar(stat = "identity")+
  ggtitle("Most common known causes of fatal collisions in NYC")+
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1, size=12),
        legend.position='none')+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 30))+
  ylab("Fatal collisions")+
  xlab(NULL)+
  scale_fill_manual(values = myColors)
p
```

# Amount of death per cause

```{r}
deaths = aggregate(collisions$NUMBER.OF.PERSONS.KILLED,
                   by=list(Category=collisions$CONTRIBUTING.FACTOR.VEHICLE.1), FUN=sum)

deaths = deaths[order(-deaths$x),]

deaths$Category = factor(deaths$Category, levels = deaths$Category)

ggplot(data=deaths[1:10,], aes(x=Category, y=x, fill=Category)) +
  geom_bar(stat = "identity")+
  ggtitle("Death per collision's cause")+
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(hjust = 1, size=8), legend.position='none')+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))+
  ylab("Deaths")
```


# Most common vehicle types involved in collisions in NYC

```{r}
# Concatenate every causes together
vehicleTypes = unlist(list(collisions$VEHICLE.TYPE.CODE.1,
                                  collisions$VEHICLE.TYPE.CODE.2, 
                                  collisions$VEHICLE.TYPE.CODE.3, 
                                  collisions$VEHICLE.TYPE.CODE.4, 
                                  collisions$VEHICLE.TYPE.CODE.5))

# Remove unknown causes
vehicleTypes = as.character(vehicleTypes)
vehicleTypes = vehicleTypes[vehicleTypes != "UNKNOWN" & vehicleTypes != ""]
vehicleTypes = as.factor(vehicleTypes)

mostCommonCauses = plyr::count(vehicleTypes) # Get frequency counts
mostCommonCauses=mostCommonCauses[order(-mostCommonCauses$freq),]

# Create dataframe with top ten causes and frequencies
vehicleTypesDataframe <- data.frame("causes"=as.character(mostCommonCauses$x[1:10]), "frequency"=
                                             mostCommonCauses$freq[1:10]) #keep top 10 only

vehicleTypesDataframe$causes = factor(vehicleTypesDataframe$causes, levels = vehicleTypesDataframe$causes)
ggplot(data=vehicleTypesDataframe, aes(x=causes, y=frequency)) + geom_bar(stat = "identity")+ggtitle("Most common vehicle type involved in collisions in NYC")+ coord_flip()+ theme(plot.title = element_text(hjust = 0.5))
```

